---
title: "Supporting Document for Thesis II"
subtitle: "Codes"
author: "Joseph"

format:
  pdf: default
  html: default

execute:
  working-dir: document

editor: visual
---

# Data Collection

## 1. Geoindicators Extraction (`gee.js`)

**Purpose:** Compute annual geospatial indicators (e.g., NDVI, NDBI, Land Surface Temperature) for each Swedish county using Google Earth Engine (GEE).

**Key Steps:**

1.  **Initialization:** Load the Earth Engine API, authenticate, and define the study region by importing county boundary shapefiles.

2.  **Satellite Band Selection:** Specify spectral bands or indices (e.g., `MODIS/006/MOD11A2` for LST, `LANDSAT/LC08/C01/T1_SR` for NDVI).

3.  **Composite Creation:** For each year:

    -   Filter image collections by date range.

    -   Apply cloud masking functions.

    -   Calculate mean or median composites per county polygon.

4.  **Indicator Computation:** Derive additional indices, e.g., NDVI = (NIR - Red) / (NIR + Red).

5.  **Aggregation:** Use `reduceRegion` with `ee.Reducer.mean()` and `ee.Reducer.stdDev()` over each county geometry.

6.  **Export:** Write the annual statistics as CSV files to a specified Google Drive folder.

```{javascript}
#| file: data/gee.js
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

## 2. Direct Unemployment Estimates (`1_SCBdirect_estimate.py`)

**Purpose:** Retrieve and process direct unemployment counts per county from Statistics Sweden (SCB) API.

**Key Steps:**

1.  **Library Imports:** Use `requests`, `pandas`, and `json` for HTTP calls and data handling.

2.  **API Authentication:** No key required; set headers for JSON content.

3.  **Endpoint Definition:** Define URL and query parameters to request unemployment by county and year.

4.  **Data Fetching:** Loop over target years, send POST requests, and collect JSON responses.

5.  **Data Cleaning:** Normalize nested JSON into tabular form, rename columns for clarity, and filter out incomplete records.

6.  **Output:** Save combined DataFrame as `direct_unemployment.csv` in the data folder.

```{python}
#| file: Python_pull/1_SCBdirect_estimate.py
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

### 3. Population Density Data (`2_SCBpopDensity.py`)

**Purpose:** Calculate population density per county by combining population counts and land area from SCB.

**Key Steps:**

1.  **Imports:** `requests`, `pandas`, and `numpy`.

2.  **Population Query:** Request `BE/BE0101` table for population totals by county.

3.  **Area Query:** Request `KM/KT0103` table for land area in square kilometers.

4.  **Merging Data:** Join population and area tables on county codes.

5.  **Density Calculation:** Compute `density = population / area` (persons per km²).

6.  **Data Validation:** Check for zero or missing areas to avoid division errors.

7.  **Output:** Export `population_density.csv` for downstream merging.

```{python}
#| file: Python_pull/2_SCBpopDensity.py
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

# Data Preprocessing

## 4. Load Libraries (`R/1_load_libraries.R`)

**Purpose:** Ensure all required R packages are installed and loaded for spatial and statistical analysis.

**Key Steps:**

1.  **Install Missing Packages:** Check for packages (`tidyverse`, `sf`, `spdep`, `INLA`, `gt`, `leaflet`, etc.) and install if absent.

2.  **Library Loading:** Load each package into the R session.

3.  **Version Logging:** Print package versions to console for reproducibility.

```{r}
#| file: R/1_load_libraries.R
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

## 5. Sweden Data Preprocessing (`R/2_sweden_preprocess.R`)

**Purpose:** Read, clean, and merge spatial and tabular data to create the analysis-ready dataset.

**Key Steps:**

1.  **Read Shapefiles:** Use `st_read` to import county boundaries (e.g., `Sweden_Counties.shp`).

2.  **Import CSVs:** Load `direct_unemployment.csv` and `population_density.csv`.

3.  **Data Cleaning:** Standardize county code formats, handle missing or NA values (e.g., impute or remove), and ensure coordinate reference systems match.

4.  **Join Datasets:** Merge shapefile with unemployment and density tables by county code.

5.  **Add Geoindicators:** Read GEE output CSVs and merge on county and year.

6.  **Final Checks:** Validate geometry integrity (`st_is_valid`), and export as an RDS object (`Sweden_data.rds`).

```{r}
#| file: R/2_sweden_preprocess.R
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

## 6. Mapping Direct Estimates (`R/3_visualization.R`)

**Purpose:** Generate both static and interactive maps to visualize direct unemployment estimates.

**Key Steps:**

1.  **Load Data:** Read `Sweden_data.rds`.

2.  **Static Maps:** Use `ggplot2`:

    -   Create choropleth layers with `geom_sf`.

    -   Customize fill scales (`scale_fill_viridis_c`), legends, and titles.

3.  **Interactive Maps:** Use `leaflet`:

    -   Convert sf object to `leaflet` object.

    -   Add polygons with `addPolygons`, tooltips for county names and values.

    -   Integrate base maps (e.g., CartoDB.Positron).

4.  **Save Outputs:** Export static maps as PNG and interactive HTML widgets.

```{r}
#| file: R/3_visualization.R
#| echo: true
#| message: false
#| warning: false
#| eval: false
```

## Small-Area Estimation

## 7. Compute Small-Area Estimates (`R/test4.R`)

**Purpose:** Fit Fay–Herriot small-area models to improve precision of county-level unemployment estimates.

**Key Steps:**

1.  **Load Data:** Read merged `Sweden_data.rds`.

2.  **Exploratory Analysis:** Compute Moran’s I (`spdep`) for spatial autocorrelation and VIF for collinearity checks.

3.  **Model Fitting:**

    -   **Untransformed Models:** Apply `lme` or `inla` for Fay–Herriot on the raw unemployment rates.

    -   **Transformed Models:** Log-transform rates, refit models, and back-transform estimates.

4.  **Diagnostics:** Plot residual vs. fitted values, Q-Q plots, and calculate MSE and CV for each area.

5.  **Result Tables:** Use `gt` to create publication-ready tables of fixed effects coefficients and error metrics.

6.  **Mapping SAE:** Add model-based estimates and CV layers to the sf object, and export final shapefile (`SAE_Results.shp`).

```{r}
#| file: R/test4.R
#| echo: true
#| message: false
#| warning: false
#| eval: false
```
